---
title: "R Notebook"
output: html_notebook
---

```{r}
library(e1071)
library(MASS)
df <- read.csv("smoking.csv")
str(df)
```

```{r}
df <- subset(df, select=-c(ID,oral))
```

Format columns.

```{r}
df$gender <- factor(df$gender)
# df$oral <- factor(df$oral) #single factor
df$dental.caries <- factor(df$dental.caries)
df$tartar <- factor(df$tartar)
df$smoking <- factor(df$smoking)
df$hearing.left. <- factor(df$hearing.left.)
df$hearing.right. <- factor(df$hearing.right.)

levels(df$dental.caries) <- c("N","Y")
levels(df$smoking) <- c("N","Y")
```

Plot smoking as a function of age.

```{r}
plot(df$smoking~df$age)
```

Plot emission as a function of population.

```{r}
plot(df$smoking~df$weight.kg.)
```

Training and testing data.

```{r}
set.seed(1234)
spec <- c(train=.6, test=.2, validate=.2)
df_ <- df
i <- sample(1:nrow(df),10000,replace=FALSE)
df <- df[i,]
i <- sample(cut(1:nrow(df),
                nrow(df)*cumsum(c(0,spec)), labels=names(spec)))
train <- df[i=="train",]
test <- df[i=="test",]
vald <- df[i=="validate",]
```

Run svm.

```{r}
svm1 <- svm(smoking~., data=train, kernel="linear", cost=10, scale=TRUE)
summary(svm1)
```

Try different costs.

```{r}
tune_svm1 <- tune(svm, smoking~., data=vald, kernel="linear", ranges=list(cost=c(0.001,0.01,0.1,1,5,10,100)))
summary(tune_svm1)
```

Try with polynomial.

```{r}
tune_svm2 <- tune(svm, smoking~., data=vald, kernel="polynomial", ranges=list(cost=c(0.001,0.01,0.1,1,5,10,100)))
summary(tune_svm2)
```

Try with radial.

```{r}
tune_svm3 <- tune(svm, smoking~., data=vald, kernel="radial", ranges=list(cost=c(0.001,0.01,0.1,1,5,10,100)))
summary(tune_svm3)
```

These algorithms are fairly slow. Each model takes about 5 minutes to calculate. This makes it difficult to make minor adjustments to the parameters for testing. It seems that radial with cost of 1 is the best. Due to the size of the dataset, I had to sample it in order to run the algorithms in a timely manner.
