---
title: "Logistic Regression from Scratch"
author: "Karen Mazidi"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

# Notebook for the Logistic Regression Chapter

First, load the package and plasma data set. 

```{r}
library(HSAUR)
df <- read.csv("titanic_project.csv")
```

### Logistic Regression using R

Use R's glm() function first as our ground truth. 


```{r}
set.seed(1234)
i <- sample(1:nrow(df), 0.75*nrow(df), replace=FALSE)
train <- df[i,]
test <- df[-i,]
glm1 <- glm(survived~sex, data=train, family=binomial)
probs <- predict(glm1,  newdata=test, type="response")
pred <- ifelse(probs> 0.5, 2, 1)
acc <- mean(pred == as.integer(test$survived))
summary(glm1)
```

### Logistic Regression from Scratch

First we need to define the sigmoid function.

```{r}
# function to return a vector of sigmoid values from an input matrix
sigmoid <- function(z){
  1.0 / (1+exp(-z))
}
# set up weight vector, label vector, and data matrix
weights <- c(1, 1)
data_matrix <- cbind(rep(1, nrow(train)), train$fibrinogen)
labels <- as.integer(train$ESR) - 1
```

Then we need code for gradient descent. The algorithm used here first starts with all weights = 1, then iterates. Notice we get the same weights (coefficients) as R gave us, but it took a lot longer.


```{r}
weights <- c(1, 1)  # repeat this for rerunning the block
learning_rate <- 0.001
for (i in 1:500000){
  prob_vector <- sigmoid(data_matrix %*% weights)
  error <- labels - prob_vector
  weights <- weights + learning_rate * t(data_matrix) %*% error
}
weights
```

### Predict with the weights we generated

```{r}
# predict with our weights
test_matrix <- cbind(rep(1, nrow(test)), test$fibrinogen)
test_labels <- as.integer(test$ESR) - 1
predicted <- test_matrix %*% weights
probabilities <- exp(predicted) / (1 + exp(predicted))
predictions <- ifelse(probabilities > 0.5, 1, 0)
mean(predictions == test_labels)
```




Visualization that the log odds is a line.

```{r}
plasma_log_odds <- cbind(rep(1, 32), df$fibrinogen) %*% weights
plot(df$fibrinogen, plasma_log_odds, col=df$ESR)
abline(weights[1], weights[2])
```